Describing demographic models is hard

### 2020-05-20 paper.tex
clean up conclusions

Fortunately, the effects of this error are subtle and it would take
and exceedingly sensitive method to detect the difference.
Population sizes and structure since the time of
the split are unaffected, so that differences in expected $F_{ST}$ are negligible between
the correct and incorrect model. However, the ancient structure distorts the distribution
of $T_{MRCA}$ beyond the split date and increases
heterozygosity in contemporary populations roughly $4\%$ over expectations
from the correct model (Fig.~\ref{fig:ooa_stats}E-F).

@@

The implementation of complex demographic models is error prone, and such errors
can have a large impact on downstream analysis and interpretation. 
The discovery and correction of the simulation errors discussed here underscore
how API design choices and 
and we recommend the following steps to ensure more robust simulations.

@@

Recognition of these difficulties drove efforts to peer-review demographic model
implementation in \stdpopsim~\citep{adrion2019community}, with quality-controlled
models and simulation resources for a growing number of commonly studied species.

The first bug discussed in this note was indeed identified through a peer-review effort of the
model implementation, and it is not easily identified by inspecting the resulting data.
The second bug would have been comparably easy to identify by inspection
of the data, but the missing parameter was easy to overlook.

We therefore recommend the following steps to ensure more robust simulations.
First, we recommend that demographic models used in publications
are verified by a second pair of eyes in code review or an independent implementation.
If the implementation is complex enough that code review is impractical, the risk
that bugs are introduced is extremely high and additional caution and statistical
verification is warranted.

Second, we recommend basic statistical verification of all simulated data.
If simulations are so large that statistical validation is burdensome,
subsets of the simulated data can be analyzed~\citep{ralph2020efficiently}.
Such tests would have caught the error in~\citet{martin2017human}.

Third, openness is essential. We only know about these errors because of open code and
open source development processes. By making their entire pipeline available,
\citet{martin2017human} not only enabled other research teams to build upon their findings,
but they made it possible for such errors to be found and corrected.
There must be many, many more mistakes out there, and we need both
pre- and post-publication vigilance from users and developers to ensure the
soundness of the large body of simulation-based analyses.

----

These errors suggests three lessons.

Second, testing of the final simulated data is important. \aprcomment{verification of simulations, wrapped together}
This can be challenging, because the amount of simulated data can be large.
However, recent progress in computing summary statistics from tree
sequence data can make this easier~\citep{ralph2020efficiently}.
Even if the entire simulation data cannot be easily evaluated, subsets of the data
can be examined to identify large errors.

Third, open data analysis pipelines are necessary for the self-correcting nature of science.
This subtle error of large effect was only discovered through our own re-use of
the simulation pipeline developed in~\citet{martin2017human} to pursue
additional analyses on a similar topic. We identified the bug by tracing down an unexpected
result in our analysis. This error could not have been found and corrected without the open
publication of the entire analysis pipeline from the original study.

>>>>

Fortunately, the effects of this error are subtle and it would take
an exceedingly sensitive method to detect the difference.
Population sizes and structure since the time of
the split are unaffected, so that differences in expected $F_{ST}$ are negligible between
the correct and incorrect model. However, the ancient structure distorts the distribution
of $T_{MRCA}$ beyond the split date and increases
heterozygosity in contemporary populations roughly $4\%$ over expectations
from the correct model (Fig.~\ref{fig:ooa_stats}E-F).

@@

The implementation of complex demographic models is error prone, and such errors
can have a large impact on downstream analysis and interpretation. 
The discovery and correction of the demographic models discussed here underscore
how API design choice can lead to the propagation of mistakes that are difficult
to notice. We therefore recommend the following steps to ensure more robust simulations.

@@

Second, from a user's perspective, steps should be taken to validate the implemented
demographic model. This could include verification through code review or an
independent implementation of the model, which is the approach taken by
\stdpopsim~\citep{adrion2019community} to compile quality-controlled
models and simulation resources for a growing number of commonly studied species.
Basic statistical validation should also be performed on any simulated model, and
recent progress in computing summary statistics from tree sequence data can make
this easier~\citep{ralph2020efficiently}. If simulations are so large that statistical
validation is burdensome, subsets of the simulated data can be analyzed to ensure
general diversity patterns are sound. Such tests likely would have caught the error
in~\citet{martin2017human}.

Finally, openness is essential to the self-correcting nature of science.
We only know about these errors because of open code and
open-source development processes. By making their entire pipeline available,
\citet{martin2017human} not only enabled other research teams to build upon their findings,
but they made it possible for such errors to be found and corrected.
There must be many, many more mistakes out there, and we need both
pre- and post-publication vigilance from users and developers to ensure the
soundness of the large body of simulation-based analyses.
